---
title: "Map Assignment"
author: "Chenghao Meng"
date: "2020/10/25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load("tidyverse","hurricaneexposuredata","drat","maps","leaflet")
```

## 1 Data Cleaning

To begin with, we will load the dataset from the package `hurricaneexposuredata`.

```{r}
# Extract data for the mapping
addRepo("geanders")
data("hurr_tracks")
data("rain")
```

Then we extract the track data of **Floyd-1999** and **Allison-2001**.

```{r}
TrackFloyd <- hurr_tracks %>% filter(storm_id=="Floyd-1999")
TrackAllison <- hurr_tracks %>% filter(storm_id=="Allison-2001")
```

After that, we will use the `filter` , `group_by` and `summarise`functions to extract rainfall data of **Floyd-1999** and **Allison-2001**.

```{r message=FALSE, warning=FALSE}
RainFloydRaw <- rain %>% 
  filter(storm_id=="Floyd-1999") %>% 
  group_by(fips,storm_id) %>% 
  summarise(sum_precip=sum(precip))

RainAllisonRaw <- rain %>% 
  filter(storm_id=="Allison-2001") %>% 
  group_by(fips,storm_id) %>% 
  summarise(sum_precip=sum(precip))
```

Since the column of fips is not that standard, we will make some transformation on that to make the all the fips numbers in a 5-digit format by using `str_pad` function in `stringr` package.

```{r}
# Make adjustments on fips
CountyFips <- county.fips
CountyFips$fips <- as.character(CountyFips$fips)
# if fips looks like "1001" rather than "01001", transformation is needed
CountyFips$fips <- str_pad(CountyFips$fips,5,side = "left",pad = "0")
```

After that, we will use `merge` function to join the CountyFips data frame into the data frames of rainfall data.

```{r}
RainFloydPre <- merge(RainFloydRaw,CountyFips,by="fips") %>% separate(polyname, into= c("region","subregion"),sep= ",")

RainAllisonPre <- merge(RainAllisonRaw,CountyFips,by="fips") %>% separate(polyname, into= c("region","subregion"),sep= ",")
```

To get the target map, we will use the `map` package to extract necessary dataset.

```{r}
MainStates <- map_data("county",c("texas","oklahoma","kansas","louisiana","arkansas","missouri","iowa","wisconsin","michigan","illinois","indiana","ohio","kentucky","tennessee","alabama","mississippi","florida","georgia","south carolina","north carolina","virginia","west virginia","maryland","delaware","pennsylvania","new jersey","new york","connecticut","rhode island","massachusetts","vermont","new hampshire","maine"))
```

Then, we will create RainFloyd and RainAllison data frames for further analysis.

```{r}
# Create RainFloyd
RainFloyd <- merge(MainStates,RainFloydPre,by=c("region","subregion"))
RainFloyd[RainFloyd==0.0] <- 0

# Make cut
RainFloyd$rain_cut <- cut(RainFloyd$sum_precip,breaks=c(0,25,50,75,100,125,150,175,200,225),ordered_result = T,include.lowest = T)
```

```{r}
# Create RainAllison
RainAllison <- merge(MainStates,RainAllisonPre,by=c("region","subregion"))
RainAllison[RainAllison==0.0] <- 0

# Make the cut
RainAllison$rain_cut <- ifelse(RainAllison$sum_precip>175,"Exposed","Unexposed")

```

Before moving on, we should test if the created colums contain NAs for the convenience of mapping.

```{r}
# NA test
anyNA(RainFloyd$rain_cut)
anyNA(RainAllison$rain_cut)
```

## 2 Map: Using ggplot

### 2.1 Floyd-1999

```{r warning=FALSE}
RainFloydPlot <- ggplot() + 
  geom_polygon(data=MainStates, aes(x=long, y=lat, group=group),colour="black",fill="white") + 
  
  geom_polygon(data=RainFloyd,aes(x=long,y=lat,group=group,fill=rain_cut),colour="transparent")+
  
  geom_path(data=TrackFloyd,aes(longitude, latitude),color="red")+
  
  xlim(min(MainStates$long),max(MainStates$long)) + 
  ylim(min(MainStates$lat),max(MainStates$lat))
```

```{r message=FALSE, warning=FALSE}
RainFloydPlot + 
  # Change the title of the legend
  labs(fill="Rainfall(mm)") + 
  xlab("Longitude") + ylab("Latitude") +
  ggtitle("Floyd-1999") + 
  scale_fill_brewer(palette="Blues") +
  theme(plot.title = element_text(hjust = 0.5))
```

### 2.2 Allison-2001

```{r}
RainAllisonPlot <- ggplot() + 
  geom_polygon(data=MainStates, aes(x=long, y=lat, group=group),colour="black",fill="white") + 
  
  geom_polygon(data=RainAllison,aes(x=long,y=lat,group=group,fill=rain_cut),colour="transparent")+
  
  geom_path(data=TrackAllison,aes(longitude, latitude),color="red")+
  
  xlim(min(MainStates$long),max(MainStates$long)) + 
  ylim(min(MainStates$lat),max(MainStates$lat))
```

```{r message=FALSE, warning=FALSE}
RainAllisonPlot + 
  # Change the title of the legend
  labs(fill="Rainfall > 175mm") +  
  xlab("Longitude") + ylab("Latitude") +
  ggtitle("Allison-2001") + 
  theme(plot.title = element_text(hjust = 0.5))
```

## 3 Map: Using leaflet

```{r}

leaflet(data = MainStates) %>% addTiles() %>%
  addPolygons(fillColor = topo.colors(10, alpha = NULL), stroke = FALSE)
```

